```{r}
library(tidyverse)
library(nflfastR)
```
# Terry McLaurin targets by half
*loading pbp data
```{r}
pbp_2022 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2022.rds")
```
22 first half
```{r}
terry_1h_22 <- pbp_2022 |>
  filter(receiver == "T.McLaurin", week<=3, game_half == "Half1")
```
23 data
```{r}
pbp_2023 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2023.rds")
```
23 first half
```{r}
terry_1h_23 <- pbp_2023 |>
  filter(receiver == "T.McLaurin", week<=3, game_half == "Half1")
```
24 data
```{r}
pbp_2024 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2024.rds") 
```
24 first half 
```{r}
terry_1h_24 <- pbp_2024 |>
  filter(receiver == "T.McLaurin", week<=3, game_half == "Half1")
```

Reflection:

Much to the displeasure of many Washington fans, the Commanders haven't frequently targeted star pass-catcher Terry McLaurin. This has been particularly apparent in the first halves of their games.

McLaurin's received just seven first-half targets through Washington's three games this year. It's a notable drop from that span last season, where the Commanders targeted the receiver 12 times. This isn't a new gripe, though — McLaurin had just six targets in the opening half of his teams first three games in 2022.

```{r}
pbp_2024 |>
  filter(!is.na(receiver)) |>
  group_by(receiver) |>
  summarize(targets = n()) |>
  arrange(desc(targets))
```

