```{r}
library(tidyverse)

```
```{r}
matches_2024 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/main/data/ncaa_womens_volleyball_matchstats_2023.csv")
```
Selecting Stats:
```{r}
team_avg <- matches_2024 |> 
  group_by(team) |> 
  summarise(
    avg_kills = mean(kills),
    avg_errors = mean(errors),
    avg_assists = mean(assists),
    avg_digs = mean(digs),
    avg_aces = mean(aces),
    avg_sets = mean(s)
    )
```
Turning stats to per set
```{r}
team_per_sets <- team_avg |>
  mutate(
    kills_set = avg_kills/avg_sets,
    errors_set = avg_errors/avg_sets,
    digs_set = avg_digs/avg_sets,
    assists_set = avg_assists/avg_sets,
    aces_set = avg_aces/avg_sets
  )



```

```{r}
per_set_zscores <- team_per_sets |> 
  mutate(
    kills_zscore = as.numeric(scale(kills_set, center = TRUE, scale = TRUE)),
    errors_zscore = as.numeric(scale(errors_set, center = TRUE, scale = TRUE)) * -1,
    digs_zscore = as.numeric(scale(digs_set, center = TRUE, scale = TRUE)),
    assists_zscore = as.numeric(scale(assists_set, center = TRUE, scale = TRUE)),
    aces_zscore = as.numeric(scale(aces_set, center = TRUE, scale = TRUE)),
    PerSetZscores = kills_zscore + errors_zscore + digs_zscore + assists_zscore + aces_zscore
  )
```
Top Z scores:
```{r}
per_set_zscores |> select(team, PerSetZscores, kills_set, errors_zscore, digs_zscore, assists_zscore, aces_zscore) |>  arrange(desc(PerSetZscores))
```

We do see a decent amount of mid-majors here, which isn't all that surprising given mid-majors typically play other mid_majors, but they're not necessarily the teams we'd expect (Dayton, Western Michigan, CSUN, Yale).

Maryland slots in at No. 279, which is one spot below Nebraska... so I clearly didn't create a very good model here!