---
title: "Women's Soccer Data"
output: html_notebook
---

We'll use this notebook to explore some NCAA women's soccer data and [our questions](https://docs.google.com/spreadsheets/d/1kTSrnl1URmcexIWLDMWqDnp3q-c2OS45uifnhjp10n4/edit#gid=0). First, let's load the tidyverse library:

```{r}
library(tidyverse)
```

We'll want to start with 2023 match data, which is located in [this GitHub repository](https://github.com/dwillis/NCAAWomensSoccerData). We can load the match data directly from GitHub using `read_csv`:

```{r}
matches_2023 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2023.csv")
```

Let's take a look at this data to see what it contains, using `head` and `glimpse`:

```{r}
head(matches_2023)
glimpse(matches_2023)
```

Now, onto the questions using aggregates, which means `group_by` and `summarize`:

1.  How many shots in total has Maryland attempted and how many have been on goal? Calculate a percentage of shots on goal.**180 shots, 77 on goal (42.8 percent)**

```{r}
maryland_2023 <- matches_2023 |>
  filter(team == "Maryland Terrapins, Terps") |>
  group_by(team)|>
  summarise(
    total_shots=sum(sh_att),
    total_sog=sum(so_g),
    sog_percentage=(total_sog/total_shots)
  )
  
```

2.  What about the same shot stats from last season's first four games? You'll need to load the matches from 2022 - just change the season in the URL - and filter by date to find the first four games. **Maryland shot better in the first four games of 2022 (47.9 percent) than it did in the 2023 season (42.8 percent). This is possibly because those four games didn't include any conference opponents, and the Big Ten competition may be more challenging than the non-conference schedule. Note: I couldn't figure out how to filter by the first four games so I used the internet, where I learned about the slice function.**

```{r}
matches_2022 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2022.csv")
```

```{r}
maryland_2022 <- matches_2022 |>
    filter(team == "Maryland Terrapins, Terps") |>
  slice(1:4)
```

```{r}
first_4_2022 <- maryland_2022 |>
  summarise(
    total_shots=sum(sh_att),
    total_sog=sum(so_g),
    sog_percentage=(total_sog/total_shots)
  ) 

```

```         
```

3.  What about draws - Maryland started 2022 with 4 ties. Have any other teams had as many this season?**Many teams had more than 4 draws in 2023. Indiana State led the country with 10, Evansville followed with 9, and 8 teams had 8.**

```{r}
draws_2023 <- matches_2023 |>
  filter(outcome == "Draw") |>
  group_by(team) |>
  summarise(
    total = n()
  ) |>
  arrange(desc(total))

```

4.  How many total corner kicks has Maryland had so far and how many have its opponents have? Calculate the difference between the two totals for every team.**Maryland took 65 corners and allowed 104. While the -39 differential feels staggering, there are three teams that took over 100 less corners than their opponents in 2023 (George Mason, Delaware St., Hampton)**

```{r}
corners_2023 <- matches_2023 |>
  group_by(team) |>
  summarise(
    corners_taken=sum(corners),
    corners_allowed=sum(defensive_corners),
    corner_differential=(corners_taken-corners_allowed)
  )
```

```{r}
matches_2024 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2024.csv")
```

```{r}
matches_2024 <- matches_2024 |>
  mutate(corner_ratio = corners/defensive_corners) |>
  filter(corner_ratio >=4)
```

```{r}
matches_2024 <- matches_2024 |>
  mutate(corner_dif=corners-defensive_corners)
```

```{r}
matches_2024 |>
    group_by(team)|>
  summarise(
    total_shots=sum(sh_att),
    total_sog=sum(so_g),
    sog_percentage=(total_sog/total_shots)
   ) |>
  arrange(desc(sog_percentage))
```
9/12 Women's Soccer Tests

```{r}
tg_23 <- matches_2023 |> 
mutate(total_goals = goals+defensive_goals)

tg_24 <- matches_2024 |>
mutate(total_goals = goals+defensive_goals)

```
T test on total goals
```{r}
result <- t.test(tg_23$total_goals, tg_24$total_goals, alternative="two.sided", var.equal = TRUE)


print(result)
```
Scoring is up this year (by about 3/4th's of a goal), and it is significant given how small the result of our p-value is. This may be because most teams haven't gone into conference play. There are far more blowouts and lopsided, high-scoring victories in non-conference play where strong teams face lesser opponents. While the averages were closer when I filtered the date, there still remained a difference of about .6 goals and the p-value remained extremely small.


What about for Maryland?
```{r}
Maryland_tg_23 <- tg_23 |> filter(team == "Maryland Terrapins, Terps") |>
mutate(total_goals = goals+defensive_goals)

Maryland_tg_24 <- tg_24 |> filter(team == "Maryland Terrapins, Terps") |>
mutate(total_goals = goals+defensive_goals)

result <- t.test(Maryland_tg_23$total_goals, Maryland_tg_24$total_goals, alternative="two.sided", var.equal = TRUE)


print(result)

```
We do see a large difference between the average total goals each year, but the p-value tells us that it is not significant. If I had to guess, I would imagine this is because of a smaller sample size. Even when I filter with the date to make it a similar number of games each season, the p-value remains high. I wonder if this is just a result of not having enough games played.


### Correlations
```{r}
correlations <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2024.csv")

```

```{r}
newcorrelations <- correlations |> 
  mutate(
    differential = goals - defensive_goals
    )
```

```{r}
newcorrelations |> summarise(correlation = cor(differential, fouls, method="pearson"))

```
```{r}
newcorrelations |> summarise(correlation = cor(differential, so_g, method="pearson"))

```
```{r}
net <- lm(differential ~ so_g, data = newcorrelations)
summary(net)
```
#Story pitch:
We know that putting shots on goals generally translates to better results, but Maryland's best scoring output in the last few years came during a game where it didn't put a ton of shots on target. So why did things go so well against VCU and how did Maryland defy such a strong trend?