```{r}
library(tidyverse)
library(dplyr)
install.packages("baseballr")
```
I had lots of trouble with my initial plan, so I decided to pivot to something else. I wanted to stay on a baseball topic because it feels like one of the more data-centric sports, and one of the most commonly used baseball data sources is baseball savant, which has become a huge deal among baseball fans. For my project, I wanted to see how much their biggest metrics (selected below) correlate to more widely-known stats like batting average, OPS, home runs, WAR, etc. — and as a result, see which of their metrics are most indicative of success.
```{r}
baseball_savant <- stats |>
  select(`last_name, first_name`, year, player_id, sweet_spot_percent, barrel_batted_rate, hard_hit_percent, avg_best_speed)
```


Now I have season recaps of 4 key savant stats from every player who's played in the last decade. Those stats:
Sweet_spot_percent: "A batted-ball event with a launch angle between eight and 32 degrees."
barrel_batted_rate:"[rate of] A batted ball with the perfect combination of exit velocity and launch angle"
hard_hit_percent: "[percentage of hits] Statcast defines a 'hard-hit ball' as one hit with an exit velocity of 95 mph or higher."
avg_best_speed: "[The] average of the hardest 50% of his batted balls."

Taking a look at the leaders in each stat:

```{r}
top_n(baseball_savant, 10, sweet_spot_percent)
```

```{r}
top_n(baseball_savant, 10, barrel_batted_rate)
```

```{r}
top_n(baseball_savant, 10, hard_hit_percent)
```

```{r}
top_n(baseball_savant, 10, avg_best_speed)
```
These top 10s give me a general idea of each of the savant stats. My next step is to join these stats with the traditional season stats so I can then figure out what the correlation is between the two.

Loading in standard stats:
```{r}
baseball_ref_24 <- read_csv("https://www.baseball-reference.com/leagues/majors/2024-standard-batting.shtml#players_standard_batting")
```

Using fangraphs as centralized source for the stats on savant and traditional stats:
```{r}
all_stats <- Fangraphs_stats_Sheet1 |>
  mutate(ab_hr = AB/HR)
```

